A standardized framework for risk-based assessment of treatment effect heterogeneity in observational healthcare databases

### 2022-06-28 submission/manuscript.rmd
update structure for resubmission

---
title: "A standardized framework for risk-based assessment of treatment effect heterogeneity in observational healthcare databases"
author: 
  - Alexandros Rekkas, MSc${^1}$ 
  - David van Klaveren, PhD$^{2,3}$, 
  - Patrick B. Ryan, PhD$^4$
  - Ewout W. Steyerberg, PhD$^{3,5}$
  - David M. Kent, PhD$^2$
  - Peter R. Rijnbeek, PhD${^1}$
output:
    bookdown::pdf_document2: default
    bookdown::word_document2:
        reference_docx: reference.docx
geometry: margin=1.0in
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left, pagewise]{lineno}
  - \usepackage{caption}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: the-lancet.csl
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(here)
library(kableExtra)
d <- function(x, decimals = 2) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}
inline_hook <- function(x) {
  if (is.numeric(x)) {
    if (abs(x - round(x)) < .Machine$double.eps) {
      formatted <- format(x, digits = 1, big.mark = " ")
    } else {
      formatted <- format(x, digits = 1, nsmall = 2, big.mark = " ", decimal.mark = intToUtf8("0x00B7"))
    }
  } else {
    formatted <- x
  }

@@

\newpage
\newpage
# Abstract {-}
\singlespacing 
**Background**: Treatment effects are often anticipated to vary across groups of
patients with different baseline risk. Implementation of a risk-based approach
in the RCT setting has shown evidence of increase in statistical power for the
evaluation of treatment effect heterogeneity. The aim of this study was to
extend this approach to the observational setting using a standardized scalable
framework.

**Methods**: The proposed framework consists of
five steps: 1) definition of the problem, i.e. the population, the treatment,
the comparator and the outcome(s) of interest; 2) identification of relevant
databases; 3) development of a prediction model for the outcome(s) of interest;
4) estimation of relative and absolute treatment effect within strata of
predicted risk, after adjusting for observed confounding; 5) presentation of the
results. We demonstrate our framework by evaluating heterogeneity of the effect
of angiotensin-converting enzyme (ACE) inhibitors versus beta blockers on three
efficacy and six safety outcomes across three observational databases.

**Findings**: In our demostration, patients at low risk of acute myocardial
infarction (MI) received negligible absolute benefits for all three efficacy
outcomes, though they were more pronounced in the highest risk quarter,
especially for hospitalization with heart failure. However, failing diagnostics
showed evidence of residual imbalances even after propensity score adjustment.

**Interpretation**: Application of our framework allows for
observing differential risk of outcomes across strata, which offers the
opportunity to consider the benefit-harm tradeoff between alternative treatments
within targeted subpopulations. Further insights may arise by application to
safety and effectiveness questions on a large scale.

**Funding**: European Health Data and Evidence Network (EHDEN) project.

@@

# Introduction

@@

# Methods

The proposed framework defines 5 distinct steps: 1) definition of the research
aim; 2) identification of the databases within which the analyses will be
performed; 3) prediction of outcomes of interest; 4) estimation of absolute and
relative treatment effects within risk strata; 5) presentation of the
results. We developed an open-source R-package for the implementation of the
proposed framework and made it publicly available
(https://github.com/OHDSI/RiskStratifiedEstimation). An overview of the entire
framework can be found in Figure \ref{fig:graphicalAbstract}.

## Case study

As a demonstration, we evaluated if our proposed method was able to identify
treatment effect heterogeneity of ACE inhibitors compared to beta blockers using
acute myocardial infarction (MI) risk quarter specific effect estimates, both on
the relative and on the absolute scale. We focused on three efficacy outcomes
(acute MI, hospitalization with heart failure and ischemic or hemorrhagic
stroke) and six safety outcomes (hypokalemia, hyperkalemia, hypotension,
angioedema, cough and abnormal weight gain). We used data from three US-based
claims databases. The analysis plan was the framework outlined in steps 1
through 5.

# Results

@@

For our demonstration we used data from three US claims databases, namely IBM
MarketScan Commercial Claims and Encounters (CCAE), IBM MarketScan Medicaid
(MDCD), and IBM MarketScan Medicare Supplemental Beneficiaries (MDCR). Our
analyses included a total of
`r patientNumbers %>% filter(database == "ccae") %>% pull(treatment)`,
`r patientNumbers %>% filter(database == "mdcd") %>% pull(treatment)`, and
`r patientNumbers %>% filter(database == "mdcr") %>% pull(treatment)`
patients initiating treatment with ACE inhibitors and 
`r patientNumbers %>% filter(database == "ccae") %>% pull(comparator)`,
`r patientNumbers %>% filter(database == "mdcd") %>% pull(comparator)`, and
`r patientNumbers %>% filter(database == "mdcr") %>% pull(comparator)`
patients initiating treatment with beta blockers in CCAE, MDCD and MDCR
respectively (Table \@ref(tab:incidenceOverall)). Adequate numbers of patients were included in all strata
of predicted acute MI risk (Supplement: Table XX).

@@

We internally developed separate prediction models for acute MI in all three
databases. The prediction models were estimated on the propensity score matched
(1:1) subset of the sample, using caliper of 
`r format(.2, nsmall = 1, decimal.mark = intToUtf8("0x00B7"))`
and after excluding patients having the outcome any time prior to treatment
initiation. We chose a 2-year time at risk for patients and developed the
prediction models using LASSO logistic regression with 3-fold cross validation
for hyper-parameter selection.

The models had moderate discriminative performance (internally validated) with
no major issues of overfitting to any cohort except for the case of CCAE, where
the derived prediction model performed better in the comparator cohort (Table
\@ref(tab:predictionPerformance)). We also observed lower performance of the
prediction model developed in MDCR compared to the other 2 databases.

@@

In the lowest acute MI risk quarter of CCAE and MDCD we observed strong
separation of the propensity score distributions, therefore, effect estimates
derived in these strata are not well-supported (Figure
\ref{fig:psDensity}). This problematic behavior is also visible in the covariate
balance plots comparing standardized mean differences of patient characteristics
before and afrer PS adjustment, where in many cases the commonly accepted bound
of 
`r format(.1, nsmall = 1, decimal.mark = intToUtf8("0x00B7"))`
is violated (Figure \ref{fig:covariateBalance}). This is more pronounced in the
lowest acute MI risk quarter of CCAE, but remains an issue for a small number of
covariates in all CCAE risk strata. This diagnostic also fails for the two lower
acute MI risk quarters of MDCD. Often the persisting imbalances were linked to
pregnancy outcomes, which can be explained by the contraindication of ACE
inhibitors in this condition. Analyses in MDCR passed all diagnostics.

Finally, the distribution of the estimated relative risks with regard to 30
negative control outcomes indicated unresolved confounding within the lowest
acute MI risk quarter of CCAE (Figrue \ref{fig:negativeControls}). Hazard ratios
significantly different than 1 (true effect size) were concentrated in the lower
right part of Figure \ref{fig:negativeControls}: panel Q1. This suggests
significant negative effects of ACE inhibitors compared to beta blockers on
causally unrelated outcomes, pointing at unresolved differences between the two
treatment arms. This was not the case in the other risk quarters of CCAE, or in
any risk quarter of MDCD and MDCR (Supplement, Figure XX).

@@

```{r, echo=FALSE}
map_outcomes <- readRDS(here("data/processed/map_outcomes.rds"))

map_exposures <- readRDS(here("data/processed/map_exposures.rds"))
overall <- readRDS(here("data/processed/mappedOverallResults.rds")) %>%
  left_join(map_outcomes, by = c("outcomeId" = "outcome_id")) %>%
  select(-outcomeId) %>%
  rename("outcome" = "outcome_name") %>%
  left_join(map_exposures, by = c("treatmentId" = "exposure_id")) %>%
  select(-treatmentId) %>%
  rename("treatment" = "exposure_name") %>%
  left_join(map_exposures, by = c("comparatorId" = "exposure_id")) %>%
  select(-comparatorId) %>%
  rename("comparator" = "exposure_name") %>%
  mutate(
    outcome = gsub("_", " ", outcome),
    estimate = paste0(
      format(round(estimate, 2), nsmall = 2, decimal.mark = intToUtf8("0x00B7")),
      " (", 
      format(round(lower, 2), nsmall = 2, decimal.mark = intToUtf8("0x00B7")),
      ", ", 
      format(round(upper, 2), nsmall = 2, decimal.mark = intToUtf8("0x00B7")),
      ")"
    )
  ) %>%
  select(-c("treatment", "comparator", "analysisType", "seLogRr", "lower", "upper")) %>%
  relocate(c("database", "outcome"))

@@

95\% CI) in CCAE and MDCD respectively (Figure 4). In MDCR hazard ratios

@@

@@

respectively (Figure 5). We found no difference on the absolute scale for stroke
across risk strata. Absolute risk differences did not favor ACE inhibitors
compared to beta blockers in terms of cough, even though this effect again
diminished with increasing acute MI risk---from 

@@

@@

# Discussion

@@

\newpage
# References
\nolinenumbers
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent
<div id="refs"></div>
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\noindent

@@

# Tables and figures

```{r incidenceOverall, echo = FALSE}
incidence <- readRDS(here::here("data/processed/incidenceOverall.rds")) %>%
  left_join(map_outcomes, by = c("outcomeId" = "outcome_id")) %>%
  select(-outcomeId) %>%
  rename("outcomeId" = "outcome_name") %>%
  mutate(
    treatmentYears = round(treatmentDays / 365),
    comparatorYears = round(comparatorDays / 365),
    outcomeId =  factor(
    outcomeId,
    levels = c(
      "acute_myocardial_infarction",
      "hospitalization_with_heart_failure",
      "stroke"
    ),
    labels = c(
        "acute myocardial infarction",
        "hospitalization with heart failure",
        "stroke"
      )
    )
  ) %>%
  arrange(database, outcomeId)

kableExtra::kbl(
  incidence %>%
    select(
      outcomeId,
      treatmentPersons,
      treatmentYears,
      treatmentOutcomes,
      comparatorPersons,
      comparatorYears,
      comparatorPersons,
      comparatorOutcomes
    ),
  align = "llrrrrr",
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = FALSE,
  caption = "Number of patients, person years and events within quarters of predicted risk for acute MI for the three efficacy outcomes of the study (acute MI, hospitalization with heart failure and ischemic or hemorrhagic stroke) across the three databases.",
  linesep = c('', '', '', '\\addlinespace'),
  col.names = c(
    "Outcome",
    "Patients",
    "Person years",
    "Outcomes",
    "Patients",
    "Person years",
    "Outcomes"
  ),
  format.args = list(
    big.mar      = " ",
    decimal.mark = "."
  )
) %>%
  kableExtra::add_header_above(
    c(
      " " = 1,
      "ACE inhibitors" = 3,
      "Beta blockers" = 3
    )
  ) %>%
  kableExtra::group_rows("CCAE", 1, 3) %>%
  kableExtra::group_rows("MDCD", 4, 6) %>%
  kableExtra::group_rows("MDCR", 7, 9) %>%
  kableExtra::kable_styling(font_size = 7)

```{r predictionPerformance, echo = FALSE}
predictionPerformance <- readRDS(here("data/processed/predictionPerformance.rds")) %>%
  filter(
    stratOutcome == 2
  ) %>%
  select(c("database", "cohort", "AUROC", "95% lower AUROC", "95% upper AUROC")) %>%
  rename(
    "lower" = "95% lower AUROC",
    "upper" = "95% upper AUROC"
  ) %>%
  mutate(
    auc = paste(
      format(round(AUROC, 2), nsmall = 2, decimal.mark = intToUtf8("0x00B7")),
      paste0(
        "(",
        format(round(lower, 2), nsmall = 2, decimal.mark = intToUtf8("0x00B7")),
        ", ",
        format(round(upper, 2), nsmall = 2, decimal.mark = intToUtf8("0x00B7")),
        ")"
      )
    )
  ) %>%
  select(-c("AUROC", "lower", "upper")) %>%
  mutate(
    cohort = factor(
      cohort,
      levels = c(
        "Matched",
        "Treatment",
        "Comparator",
        "EntirePopulation"
      ),
      labels = c(
        "Matched",
        "Treatment",
        "Comparator",
        "Entire population"
      )
    ),
    database = factor(
      database,
      levels = c(
        "ccae",
        "mdcd",
        "mdcr"
      )
    )
  ) %>%
  arrange(database, cohort)

```{r overallTable, echo = FALSE}
kableExtra::kbl(
  overall %>% 
    pivot_wider(id_cols = "outcome", names_from = "database", values_from = "estimate"),
  format = "latex",
  align = "lrrr",
  caption = "Overall hazard ratios.",
  booktabs = TRUE,
  col.names = c(
    "Outcome",
    "CCAE",
    "MDCD",
    "MDCR"
  ),
  format.args = list(decimal.mark = intToUtf8("0x00B7"))
) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r graphicalAbstract, cache=TRUE, echo=FALSE, fig.cap="Illustration of how the framework is applied on two observational databases, preferably mapped to OMOP-CDM.", fig.show="hold", out.width = '100%'}
include_graphics(here("figures/graphicalAbstract.png"))
```

```{r psDensity, cache=TRUE, echo=FALSE, fig.cap="Preference score distributions for the evaluation of heterogeneity of the effect of ACE inhibitors compared to beta blockers on acute MI based on quarters of predicted acute MI risk. The preference score is a transformation of the propensity score that adjusts for prevalence differences between populations.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/PsDensity.tiff")))
```

```{r covariateBalance, cache=TRUE, echo=FALSE, fig.cap="Patient characteristic balance for ACE inhibitors and beta blockers before and after stratification on the propensity scores. Each dot represents the standardized difference of means for a single covariate before (x-axis) and after (y-axis) stratification.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/CovariateBalance.tiff")))
```

```{r negativeControls, cache=TRUE, echo=FALSE, fig.cap=" Systematic error. Effect size estimates for the negative controls (true hazard ratio = 1). Estimates below the diagonal dashed lines are statistically significant (alpha = 0.05) different from the true effect size. A well-calibrated estimator should have the true effect size within the 95 percent confidence interval 95 percent of times.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/NegativeControlsPlot_ccae.tiff")))
```

```{r relative, cache=TRUE, echo=FALSE, fig.cap="Overview of heterogeneity of ACE inhibitors treatment on the relative scale (hazard ratios) within strata of predicted risk of acute MI. Values below 1 favor ACE inhibitors, while values above 1 favor beta blockers.", fig.show="hold", out.width = '100%'}
    grid::grid.raster(tiff::readTIFF(here("figures/CombinedRelative_acute_myocardial_infarction.tiff")))
```

```{r absolute, cache=TRUE, echo=FALSE, fig.cap="Overview of heterogeneity of ACE inhibitors treatment on the absolute scale within strata of predicted risk of acute MI. Estimates of absolute treatment effect are derived as the difference in Kaplan-Meier estimates at two years after inclusion. Values above 0 favor ACE inhibitors, while values below 0 favor beta blockers.  ", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/CombinedAbsolute_acute_myocardial_infarction.tiff")))
```

>>>>

---
title: "A standardized framework for risk-based assessment of treatment effect heterogeneity in observational healthcare databases"
author: 
  - Alexandros Rekkas, MSc${^1}$ 
  - David van Klaveren, PhD$^{2,3}$, 
  - Patrick B. Ryan, PhD$^4$
  - Ewout W. Steyerberg, PhD$^{3,5}$
  - David M. Kent, MD$^2$
  - Peter R. Rijnbeek, PhD${^1}$
output:
    bookdown::pdf_document2:
      number_sections: false
    bookdown::word_document2:
        reference_docx: reference.docx
geometry: margin=1.0in
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left, pagewise]{lineno}
  - \usepackage[labelfont=bf]{caption}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: the-lancet.csl
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(here)
library(kableExtra)
d <- function(x, decimals = 2) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}
inline_hook <- function(x) {
  if (is.numeric(x)) {
    if (abs(x - round(x)) < .Machine$double.eps) {
      formatted <- format(x, digits = 1, big.mark = ",")
    } else {
      formatted <- format(x, digits = 1, nsmall = 2, big.mark = ",", decimal.mark = ".")
    }
  } else {
    formatted <- x
  }

@@

\newpage
\newpage
# Abstract {-}
\singlespacing 
Treatment effects are often anticipated to vary across groups of patients with
different baseline risk. The Predictive Approaches to Treatment Effect
Heterogeneity (PATH) statement focused on baseline risk as a robust predictor of
treatment effect and provided guidance on risk-based assessment of treatment
effect heterogeneity in the RCT setting. The aim of this study was to extend
this approach to the observational setting using a standardised scalable
framework. The proposed framework consists of five steps: 1) definition of the
research aim, i.e., the population, the treatment, the comparator and the
outcome(s) of interest; 2) identification of relevant databases; 3) development
of a prediction model for the outcome(s) of interest; 4) estimation of relative
and absolute treatment effect within strata of predicted risk, after adjusting
for observed confounding; 5) presentation of the results. We demonstrate our
framework by evaluating heterogeneity of the effect of angiotensin-converting
enzyme (ACE) inhibitors versus beta blockers on three efficacy and six safety
outcomes across three observational databases.  The proposed framework can
supplement any comparative effectiveness study. We provide a publicly available
R software package for applying this framework to any database mapped to the
Observational Medical Outcomes Partnership Common Data Model. In our
demonstration, patients at low risk of acute myocardial infarction received
negligible absolute benefits for all three efficacy outcomes, though they were
more pronounced in the highest risk quarter, especially for hospitalisation with
heart failure. However, failing diagnostics showed evidence of residual
imbalances even after adjustment for observed confounding. Our framework allows
for the evaluation of differential treatment effects across risk strata, which
offers the opportunity to consider the benefit-harm trade-off between
alternative treatments. It is easily applicable and highly informative whenever
treatment effect estimation in observational data is of interest.

@@

# INTRODUCTION

@@

# RESULTS
The proposed framework defines 5 distinct steps: 1) definition of the research
aim; 2) identification of the databases within which the analyses will be
performed; 3) prediction of outcomes of interest; 4) estimation of absolute and
relative treatment effects within risk strata; 5) presentation of the
results. We developed an open-source R-package for the implementation of the
proposed framework and made it publicly available
(https://github.com/OHDSI/RiskStratifiedEstimation). An overview of the entire
framework can be found in Figure \ref{fig:graphicalAbstract}.

```{r graphicalAbstract, cache=TRUE, echo=FALSE, fig.cap="Illustration of how the framework is applied on two observational databases, preferably mapped to OMOP-CDM.", fig.show="hold", out.width = '100%'}
include_graphics(here("figures/rsee_process.png"))
```

As a demonstration, we evaluated if our proposed method was able to identify
treatment effect heterogeneity of ACE inhibitors compared to beta blockers using
acute myocardial infarction (MI) risk quarter specific effect estimates, both on
the relative and on the absolute scale. We focused on three efficacy outcomes
(acute MI, hospitalization with heart failure and ischemic or hemorrhagic
stroke) and six safety outcomes (hypokalemia, hyperkalemia, hypotension,
angioedema, cough and abnormal weight gain). We used data from three US-based
claims databases.

@@

For our demonstration we used data from three US claims databases, namely IBM
MarketScan Commercial Claims and Encounters (CCAE), IBM MarketScan Medicaid
(MDCD), and IBM MarketScan Medicare Supplemental Beneficiaries (MDCR). Our
analyses included a total of
`r patientNumbers %>% filter(database == "ccae") %>% pull(treatment)`,
`r patientNumbers %>% filter(database == "mdcd") %>% pull(treatment)`, and
`r patientNumbers %>% filter(database == "mdcr") %>% pull(treatment)`
patients initiating treatment with ACE inhibitors and 
`r patientNumbers %>% filter(database == "ccae") %>% pull(comparator)`,
`r patientNumbers %>% filter(database == "mdcd") %>% pull(comparator)`, and
`r patientNumbers %>% filter(database == "mdcr") %>% pull(comparator)`
patients initiating treatment with beta blockers in CCAE, MDCD and MDCR
respectively (Table \@ref(tab:incidenceOverall)). Adequate numbers of patients
were included in all strata of predicted acute MI risk (Supplement, page 26).

```{r incidenceOverall, echo = FALSE}
map_outcomes <- readRDS(here("data/processed/map_outcomes.rds"))
map_exposures <- readRDS(here("data/processed/map_exposures.rds"))

incidence <- readRDS(here::here("data/processed/incidenceOverall.rds")) %>%
  left_join(map_outcomes, by = c("outcomeId" = "outcome_id")) %>%
  select(-outcomeId) %>%
  rename("outcomeId" = "outcome_name") %>%
  mutate(
    treatmentYears = round(treatmentDays / 365),
    comparatorYears = round(comparatorDays / 365),
    outcomeId =  factor(
    outcomeId,
    levels = c(
      "acute_myocardial_infarction",
      "hospitalization_with_heart_failure",
      "stroke"
    ),
    labels = c(
        "acute myocardial infarction",
        "hospitalization with heart failure",
        "stroke"
      )
    )
  ) %>%
  arrange(database, outcomeId)

kableExtra::kbl(
  incidence %>%
    select(
      outcomeId,
      treatmentPersons,
      treatmentYears,
      treatmentOutcomes,
      comparatorPersons,
      comparatorYears,
      comparatorPersons,
      comparatorOutcomes
    ),
  align = "llrrrrr",
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = FALSE,
  caption = "Number of patients, person years and events within quarters of predicted risk for acute MI for the three efficacy outcomes of the study (acute MI, hospitalization with heart failure and ischemic or hemorrhagic stroke) across the three databases.",
  linesep = c('', '', '', '\\addlinespace'),
  col.names = c(
    "Outcome",
    "Patients",
    "Person years",
    "Outcomes",
    "Patients",
    "Person years",
    "Outcomes"
  ),
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::add_header_above(
    c(
      " " = 1,
      "ACE inhibitors" = 3,
      "Beta blockers" = 3
    )
  ) %>%
  kableExtra::group_rows("CCAE", 1, 3) %>%
  kableExtra::group_rows("MDCD", 4, 6) %>%
  kableExtra::group_rows("MDCR", 7, 9) %>%
  kableExtra::kable_styling(font_size = 7)

@@

We internally developed separate prediction models for acute MI in all three
databases. The prediction models were estimated on the propensity score matched
(1:1) subset of the sample, using caliper of 
`r format(.2, nsmall = 1, decimal.mark = ".")`
and after excluding patients having the outcome any time prior to treatment
initiation. We chose a 2-year time at risk for patients and developed the
prediction models using LASSO logistic regression with 3-fold cross validation
for hyper-parameter selection.

The models had moderate discriminative performance (internally validated) with
no major issues of overfitting to any cohort except for the case of CCAE, where
the derived prediction model performed better in the comparator cohort (Table
\@ref(tab:predictionPerformance)). We also observed lower performance of the
prediction model developed in MDCR compared to the other 2 databases
(Supplement, page 29).

```{r predictionPerformance, echo = FALSE}
predictionPerformance <- readRDS(here("data/processed/predictionPerformance.rds")) %>%
  filter(
    stratOutcome == 2
  ) %>%
  select(c("database", "cohort", "AUROC", "95% lower AUROC", "95% upper AUROC")) %>%
  rename(
    "lower" = "95% lower AUROC",
    "upper" = "95% upper AUROC"
  ) %>%
  mutate(
    auc = paste(
      format(round(AUROC, 2), nsmall = 2), 
      paste0(
        "(",
        format(round(lower, 2), nsmall = 2),
        ", ",
        format(round(upper, 2), nsmall = 2),
        ")"
      )
    )
  ) %>%
  select(-c("AUROC", "lower", "upper")) %>%
  mutate(
    cohort = factor(
      cohort,
      levels = c(
        "Matched",
        "Treatment",
        "Comparator",
        "EntirePopulation"
      ),
      labels = c(
        "Matched",
        "Treatment",
        "Comparator",
        "Entire population"
      )
    ),
    database = factor(
      database,
      levels = c(
        "ccae",
        "mdcd",
        "mdcr"
      )
    )
  ) %>%
  arrange(database, cohort)

@@

In the lowest acute MI risk quarter of CCAE and MDCD we observed strong
separation of the propensity score distributions, therefore, effect estimates
derived in these strata are not well-supported (Figure
\ref{fig:psDensity}). This problematic behavior is also visible in the covariate
balance plots comparing standardized mean differences of patient characteristics
before and afrer PS adjustment, where in many cases the commonly accepted bound
of 
`r format(.1, nsmall = 1, decimal.mark = ".")`
is violated (Figure \ref{fig:covariateBalance}). This is more pronounced in the
lowest acute MI risk quarter of CCAE, but remains an issue for a small number of
covariates in all CCAE risk strata. This diagnostic also fails for the two lower
acute MI risk quarters of MDCD. Often the persisting imbalances were linked to
pregnancy outcomes, which can be explained by the contraindication of ACE
inhibitors in this condition. Analyses in MDCR passed all diagnostics.

```{r psDensity, cache=TRUE, echo=FALSE, fig.cap="\\textbf{Preference score distributions in quarters of predicted acute MI risk}. The preference score is a transformation of the propensity score that adjusts for prevalence differences between populations. Higher overlap of the preference score distributions indicates that patients in the target and the comparator cohorts are more similar in terms of the predicted probability of receiving treatment (ACE inhibitors).", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/PsDensity.tiff")))
```

```{r covariateBalance, cache=TRUE, echo=FALSE, fig.cap="\\textbf{Patient characteristic balance for ACE inhibitors and beta blockers before and after stratification on the propensity scores}. Each dot represents the standardized difference of means for a single covariate before (x-axis) and after (y-axis) stratification. A commonly used rule of thumb suggests that standardized mean differences above 0.1 indicate insufficient covariate balance post propensity score adjustment.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/CovariateBalance.tiff")))
```

Finally, the distribution of the estimated relative risks with regard to 30
negative control outcomes indicated unresolved confounding within the lowest
acute MI risk quarter of CCAE (Figrue \ref{fig:negativeControls}). Hazard ratios
significantly different than 1 (true effect size) were concentrated in the lower
right part of Figure \ref{fig:negativeControls}: panel Q1. This suggests
significant negative effects of ACE inhibitors compared to beta blockers on
causally unrelated outcomes, pointing at unresolved differences between the two
treatment arms. This was not the case in the other risk quarters of CCAE, or in
any risk quarter of MDCD and MDCR (Supplement, pages 34-36).

```{r negativeControls, cache=TRUE, echo=FALSE, fig.cap="\\textbf{Systematic error}. Effect size estimates for the negative controls (true hazard ratio = 1). Estimates below the diagonal dashed lines are statistically significant (alpha = 0.05) different from the true effect size. A well-calibrated estimator should have the true effect size within the 95 percent confidence interval 95 percent of times.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/NegativeControlsPlot_ccae.tiff")))
```

@@

```{r, echo=FALSE}
overall <- readRDS(here("data/processed/mappedOverallResults.rds")) %>%
  left_join(map_outcomes, by = c("outcomeId" = "outcome_id")) %>%
  select(-outcomeId) %>%
  rename("outcome" = "outcome_name") %>%
  left_join(map_exposures, by = c("treatmentId" = "exposure_id")) %>%
  select(-treatmentId) %>%
  rename("treatment" = "exposure_name") %>%
  left_join(map_exposures, by = c("comparatorId" = "exposure_id")) %>%
  select(-comparatorId) %>%
  rename("comparator" = "exposure_name") %>%
  mutate(
    outcome = gsub("_", " ", outcome),
    estimate = paste0(
      format(round(estimate, 2), nsmall = 2),
      " (", 
      format(round(lower, 2), nsmall = 2),
      ", ", 
      format(round(upper, 2), nsmall = 2),
      ")"
    )
  ) %>%
  select(-c("treatment", "comparator", "analysisType", "seLogRr", "lower", "upper")) %>%
  relocate(c("database", "outcome"))

@@

95\% CI) in CCAE and MDCD respectively (Figure \ref{fig:relative}). In MDCR hazard ratios

@@

```{r overallTable, echo = FALSE}
kableExtra::kbl(
  overall %>% 
    pivot_wider(id_cols = "outcome", names_from = "database", values_from = "estimate"),
  format = "latex",
  align = "lrrr",
  caption = "Overall hazard ratios.",
  booktabs = TRUE,
  col.names = c(
    "Outcome",
    "CCAE",
    "MDCD",
    "MDCR"
  )) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r relative, cache=TRUE, echo=FALSE, fig.cap="\\textbf{Overview of heterogeneity of ACE inhibitors treatment on the relative scale (hazard ratios) within strata of predicted risk of acute MI}. Values below 1 favor ACE inhibitors, while values above 1 favor beta blockers.", fig.show="hold", out.width = '100%'}
    grid::grid.raster(tiff::readTIFF(here("figures/CombinedRelative_acute_myocardial_infarction.tiff")))
```

@@

respectively (Figure \ref{fig:absolute}). We found no difference on the absolute
scale for stroke across risk strata. Absolute risk differences did not favor ACE
inhibitors compared to beta blockers in terms of cough, even though this effect
again diminished with increasing acute MI risk---from

@@

```{r absolute, cache=TRUE, echo=FALSE, fig.cap="\\textbf{Overview of heterogeneity of ACE inhibitors treatment on the absolute scale within strata of predicted risk of acute MI}. Estimates of absolute treatment effect are derived as the difference in Kaplan-Meier estimates at two years after inclusion. Values above 0 favor ACE inhibitors, while values below 0 favor beta blockers.  ", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/CombinedAbsolute_acute_myocardial_infarction.tiff")))
```

@@

# DISCUSSION

@@

# METHODS

# DATA AVAILABILITY
The claims data are proprietary and are not publicly accessible due to
restricted user agreement. Database descriptions are available in the
Supplementary Material.

# CODE AVAILABILITY
Our analysis code is publicly available under the MIT license at
https://github.com/mi-erasmusmc/AceBeta9Outcomes.

\newpage
# REFERENCES
\nolinenumbers
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent
<div id="refs"></div>
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\noindent

@@

# AUTHOR CONTRIBUTIONS
AR, PRR, and DVK conceptualised the study. AR, DVK, EWS, and DMK developed the
methodology. PBR, and PRR acquired the data and AR analysed the data. AR
developed the software and wrote drafted the manuscript, which was critically
reviewed by DVK, PBR, EWS, DMK, and PRR. AR, PBR, and PRR had full access to the
raw data. All authors read and approved the manuscript and had final
responsibility for the decision to submit for publication.

# COMPETING INTERESTS
AR and PRR work for a group that received unconditional research grants from
Boehringer-Ingelheim, GSK, Janssen Research & Development, Novartis, Pfizer,
Yamanouchi, and Servier. None of these grants result in a conflict of interest
for the content of this paper. PBR is an employee of Janssen R&D, subsidiary of
Johnson & Johnson. DVK, DMK, EWS have nothing to declare.
